---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r read data, echo=TRUE, results = "asis"}
data <- read.csv(unz("activity.zip","activity.csv"),header=TRUE, stringsAsFactors = FALSE)
str(data)
data$date <- as.Date(data$date,format="%Y-%m-%d")
data$interval <- as.factor(data$interval)
str(data)
```

## What is mean total number of steps taken per day? - ignore missing values
Calculate the total number of steps taken per day
Make a histogram of the total number of steps taken each day
Mean and median of the total number of steps taken per day

```{r total steps, echo=TRUE, results = "asis"}
stepspd <- tapply(data$steps,data$date,sum)
hist(stepspd)
summary(stepspd)
means <- mean(stepspd, na.rm = T)
medians <- median(stepspd, na.rm = T)
```

The mean of total number of steps is `r means` and 
median is `r medians`.

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r pattern, echo=TRUE, results = "asis"}
p <- tapply(data$steps,data$interval,mean,na.rm = T)
plot(p, type = "l", main = "Average steps across all days", 
     xlab = "Interval", ylab = "Average steps")
maxint <- row.names(p)[which.max(p)]
```

The 5-minute interval which contains the maximum number of steps is 
the interval: `r maxint`

## Imputing missing values



Impute missing values
<http://www.mail-archive.com/r-help@r-project.org/msg58289.html>

```{r impute_missing, echo=TRUE, results = "asis"}
narows <- nrow(data)-nrow(data[complete.cases(data),])

stepsna <- is.na(data$steps)
datena <- is.na(data$date)
intna <- is.na(data$interval)
sum(datena) # no missing values in date
sum(intna) # no missing values in interval

library(plyr)
impute <- function(x, fun) {
  missing <- is.na(x)
  replace(x, missing, fun(x[!missing]))
}
# impute missing in steps as mean for a day
data2 <- ddply(data, ~ interval, transform, steps = impute(steps, mean))
sum(is.na(data$steps))
stepspd2 <- tapply(data2$steps,data2$date,sum)
hist(stepspd2)
summary(stepspd2)
means2 <- mean(stepspd2, na.rm = T)
medians2 <- median(stepspd2, na.rm = T)

```

The total number of rows with missing values is: `r narows`

The mean of total number of steps is `r means2` and 
median is `r medians2`.

As a result of imputing missing values the median has incrased but the mean remains the same.

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r DayofWeek, echo=TRUE, results = "asis"}

#create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#Use `%in%` and `weekdays` to create a logical vector
#convert to `factor` and specify the `levels/labels`
data2$wDay <- factor((weekdays(data2$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
         
```        
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r DayofWeekplot, echo=TRUE, results = "asis"}

p <- aggregate(steps ~ wDay+interval, data= data2, mean)
library(lattice)
xyplot(steps~interval|wDay, type = "l", data = p, 
     xlab = "Interval", ylab = "Number of steps",layout=c(1,2))


```